---
title: "10X/BD data integration for healthy samples"
author: "Isaac Garcia-Salinas"
date: "30/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Here I summarize the main results from my rotation project, which main goal was to integrate single cell multimodal data (targeted mRNA + protein counts)
generated using Becton-Dickinson (BD) Rhapsody technology, and a 10X (3') whole transcriptome sequencing data. Both datasets consist of CD4+ T cells isolated 
from peripheral blood mononuclear cells from healthy donors, one in the BD Rhapsody dataset, and two in the 10X dataset. For integration, I used the 
scvi-tools python library (Lopez et al., 2018), and their multimodal data analysis pipeline called totalVI (Gayoso et al., 2020). Briefly, this method aims 
to model a cell’s state based on a joint analysis of protein and RNA data. Further, it allows to remove batch-effects and impute protein expression values 
to cells missing this information level. For this, totalVI loads transcriptome and protein counts, as well as covariate level information, into an 
encoder neural network that is used to build a low dimensional representation of a cell's state (Gayoso et al., 2020).

CD4+ T cell data generated with the BD technology was available as part of the Trzupek et al., 2019 publication, which included targeted transcriptome 
sequencing of 399 genes, and profiling for up to 56 cell surface proteins. This dataset was accessed with the GEO accession series ID GSE150060, from where I obtained
counts matrices and cell-level metadata. Whole 10X data was produced by the Trynka team as part of the immunodeficiencies project. 

In the next chunks, I describe the main methods I followed to integrate both datasets. All code shown in this markdown should be run while maintaining the 
directory structure set in the "igs_rotation" drive folder.

```{r all_setup, echo=FALSE}

set.seed(123456)

lapply(c("dplyr", "Seurat", "ggplot2", "tidyverse", "ggsci", "rlang", 
         "ggpubr"), library, character.only=T)

main_dir <- getwd()  
data_dir <- paste0(main_dir, "/data/")
BD_data_dir <- paste0(data_dir, "paper_original_datasets/")
write_dir <- paste0(main_dir, "/write/")
scripts_dir <- paste0(main_dir, "/scripts/")
#plots_dir <- paste0(main_dir, "/plots/")

lapply(c("loadSingleDataset_Ferreira2020_data_spec.R",
         "ggplot_final_Processing.R", 
         "expressionHeatmaps.R",  
         "scPlot_themes.R"), function(aScript){
           loadThis <- paste0(scripts_dir, aScript)
           source(file = loadThis, local = F)
         })


```

## Reading data

In the next chunk I use a function that reads csv BD data, while subseting to non-multiplet cells, as identified in each csv metadata. As the 
Trzupek, et al., 2019 data contains three donors in total (one healthy, two non-healthy), it is possible to read their information with the 
same function. Note: Only resting cell information can be retrieved.


```{r reading_BD_data, eval=FALSE}

# Next function can also return a matrix instead of a seurat by setting 
# returnSeurat to FALSE. In addtion, dataType can be "all", "rna", or "protein"
# to retrieve data type subsets of a given donor. In all cases multiplet cells are 
# removed

healthy_seu_raw <- loadSingleDataset(dataDir = BD_data_dir, loadWhat = "healthy", 
                                     dataType = "all", returnSeurat = TRUE)
t1d_seu_raw <- loadSingleDataset(dataDir = BD_data_dir, loadWhat = "t1d", 
                                 dataType = "all", returnSeurat = TRUE)
sle_seu_raw <- loadSingleDataset(dataDir = BD_data_dir, loadWhat = "sle", 
                                 dataType = "all", returnSeurat = TRUE)


```


# Features in the BD datasets

As described in the original paper, some of the previous datasets profiled a slightly different number features, out of a 
total of 399 genes and 56 proteins for the datasets altogether (455 total features). For instance, the  SLE dataset contains 436 features, 
and 13 features are unique to this dataset. In the case of the Healthy and T1D dataset, the same number of features were profiled, and total to 
442 features. From these, 19 are not present in the SLE dataset. 


```{r shared features between the Trzupek 2019 datasets, echo=FALSE} 

print(  paste0("SLE data total features: ", length(rownames(sle_seu_raw)) ) ) 

print(  paste0("T1D data total features: ", length(rownames(t1d_seu_raw)) ) ) 

print(  paste0("Healthy data total features: ", length(rownames(healthy_seu_raw)) ) ) 

print( paste0("Intersecting features between T1D and Healthy datasets: ", 
              length(intersect(rownames(healthy_seu_raw), rownames(t1d_seu_raw)) ) )   )

print( paste0("Intersecting features between T1D/Healthy and SLE datasets: ", 
              length( intersect(rownames(sle_seu_raw), rownames(t1d_seu_raw)) ) )   )

print( paste0("Features unique to the SLE dataset: ", 
              length( setdiff(rownames(sle_seu_raw), rownames(t1d_seu_raw)) ) )   )

print( paste0("Features unique to the T1D/Healthy dataset: ", 
              length( setdiff(rownames(t1d_seu_raw), rownames(sle_seu_raw)) ) )   )

print( paste0("Total features (all datasets union) ", 
              length( unique(c(rownames(t1d_seu_raw), rownames(sle_seu_raw), rownames(healthy_seu_raw)) ) ) ))

# allDatasetsFeatVenn <- venn.diagram(x=list(rownames(t1d_seu_raw), rownames(healthy_seu_raw), 
# rownames(sle_seu_raw) ), category.names = c("T1D" , "Healthy", "SLE"),filename = NULL)
# grid.draw(allDatasetsFeatVenn)

```


Prior integrating 10X and BD data, I tested the integration of all BD datasets as means to reproduce some of 
the main findings in the Trzupek paper, and to standardize the analysis of multimodal data. As part of this, I 
merged such datasets, and because of this operation, counts for features not profiled in a given sample were 
set to 0 in the cells in that sample. For this reason, further analyses using counts for the healthy sample only, 
show total feature counts that reflect the union of all features in all the datasets reported by Trzupek et al., 
(i.e. 455 features, 399 genes and 56 proteins). 

As part of this repository data, I have included the healthy counts matrix reflecting the processing described 
above. Features in this matrix are further arranged to first show transcript features, and then protein features, 
which will be helpful when converting this data to the h5 format, which is required for totalVI-based analyses. 
This counts matrix can be found under the name "bd_healthy_counts_matrix.csv".

The feature arrangement in the counts matrix described above follows the row order shown in the 
"BD features table", where I summarize all the features profiled in the Trzupek publication. This resource is also 
available in the data directory under the name "allBD_features_table.txt".


## Recovering healthy samples counts matrix from 10X datatsets 

In the next chunk, I show the processing of the 10X data to be integrated with the BD healthy dataset.
For this, I merge each of the 10X healthy datasets using Seurat, and recover the counts matrix of the final object.
To be able to impute proteins using totalVI, I generated a 0-filled matrix for protein features that is appended to 
the 10X matrix.


```{r recovering the 10X counts matrix, eval=FALSE}

# 3 seurat objects contained transcript counts for 2 healthy donors 
# in the immunodeficiencies project

# Next is a table containing all features analyzed in the BD datasets
allBD_features_ordered <-  read.csv(file = paste0(data_dir, "allBD_features_table.tsv"), 
                                   sep="\t", header = T)
# Select protein features to build a zero-filled matrix
allBDProtFeatures <- subset(allBD_features_ordered, suffix_dtype == "pAbO")[["gene.prot_name"]]


p1 <- readRDS(paste0(data_dir, "PID_10X7838125_Ctrl_1.CD4.rds"))
DefaultAssay(p1) <- "RNA"
p2 <- readRDS(paste0(data_dir, "PID_10X8012526_CTRL_3.CD4.rds"))
DefaultAssay(p2) <- "RNA"
p3 <- readRDS(paste0(data_dir, "PID_10X8012527_CTRL_3.CD4.rds"))
DefaultAssay(p3) <- "RNA"

pats_temp <- merge(x= p1, y = p2)
allControls <- merge(pats_temp, p3)

# Adding protein features to 10X data as zeroes
sim_missingProteinCounts_10X <- matrix(rep(0,285208), 
                                       nrow=length(allBDProtFeatures), 
                                       ncol =  length(colnames(allControls)), 
                                       dimnames = list(allBDProtFeatures, 
                                                       colnames(allControls))) 
# recover 10X matrix for merged datasets
allcontrolsMatrix <- as.matrix(GetAssayData( object =allControls, slot="counts", assay="RNA"))

# append zeroes-matrix and export as csv
write.table(x = rbind(as.matrix(allcontrolsMatrix, sim_missingProteinCounts_10X), 
            file = paste0(write_dir, "three_healthy10X_counts_matrix_wMissingProt.csv"), 
            sep = ",", quote = F, row.names = T, col.names = T)

```


## Converting csv counts matrices to h5 matrices

The totalVI pipeline is part of the scvi-tools python library, which require to read mutimodal data formatted as an h5
matrix. Although several tools exist to convert directly between Seurat (or csv) format to h5, none of this allows to 
keep metadata to identify the protein features, which is required for further processing using scvi-tools. For this reason, 
I wrote a python script that allowed me to convert a csv counts matrix to an h5 file that was able to solver this problem
This mimics the format of an h5 matrix containing multimodal data as if it was natively
produced with the 10X processing software. In the next chunk I provide an example of how to execute this script. I include
the h5 matrices obtained with this method as part of this repository data folder.


```{bash Converting csv to h5, eval=FALSE, engine="sh"}

# headerless_bd_geneNames.txt correspond to a txt with a feature name per row
# The order of this file has to follow the order of the matrix features

currentDir=$(pwd)
data_folder=${currentDir}"/data"
bd_feature_names="headerless_bd_geneNames.txt"
bd_gene_count=399
bd_protein_count=56
bd_healthy_csv="bd_healthy_counts_matrix.csv"


# Convert csv files to h5 matrices
./csv2h5_igs.py -g ${bd_gene_count} -p ${bd_protein_count} -m ${bd_healthy_csv} -d ${bd_feature_names} -s ${data_folder} -o bd_healthy_counts

```


## Integrating BD and 10X data using scvi-tools and totalVI

h5 matrices obtained with the above method are used for integration in scvi-tools. 
As explained in this' package documentation (https://www.scvi-tools.org/en/stable/installation.html), 
a specific conda environment is required for execution. In the next chunk, I include the code used to integrate 
the healthy samples BD and 10X datasets.


```{python BD/10X integration using scvi-tools and totalVI, eval=FALSE}

import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import os 
import scvi

import scanpy as sc

currDir = os.getcwd()
data_dir = currDir+"/data/"
write_dir = currDir+"/write/"

# Read data using scanpy
healthy_10x = sc.read_10x_h5(filename=data_dir+"three_healthy10X_counts_matrix_wMissingProt.h5",
gex_only = False)
healthy_bd = sc.read_10x_h5(filename=data_dir+"bd_healthy_counts_matrix.h5", 
gex_only = False)

# Metadata containing total number of features per cell and sample origin
all_metadata = pd.read_csv(data_dir+"all_10xBD_metadata.csv", delimiter=",", header=0)
# Supplied metadata contains all samples (t1d/sle/heathy for BD data, and the 2 controls from 10X)
# In here, I select only metadata corresponding to the healthy samples - "Ctrl_1","CTRL_3" correspond
# to 10X data 
all_metadata=all_metadata.loc[all_metadata['sample'].isin(["Ctrl_1","CTRL_3", "Healthy"])]


# Concatenate anndata objects using scanpy
alldata_sc = healthy_10x.concatenate(healthy_bd,index_unique=None)

# add metadata 
alldata_sc.obs["nFeature_RNA"] = all_metadata["nFeature_RNA"]
alldata_sc.obs["sample"] = all_metadata["sample"]
alldata_sc.obs["technology"] = all_metadata["technology"]

# next scvi-tools command differentiates between protein and transcript data
# this information is available as part of the h5 matrices
# this command puts protein expression in alldata_sc.obms["protein_expression]
scvi.data.organize_cite_seq_10x(alldata_sc) 


# Scvi-tools models use raw count data, but normalization can be performed for other 
# applications, just ensure that raw counts are kept unchanged
alldata_sc.layers["counts"] = alldata_sc.X.copy()
sc.pp.normalize_total(alldata_sc, target_sum=1e4)
sc.pp.log1p(alldata_sc)
alldata_sc.raw = alldata_sc

# totalVI model runs over variable features, that can be calculated using standard 
# seurat methods, in this case, the variance stabilization method (vst). 
sc.pp.highly_variable_genes(
    alldata_sc,
    n_top_genes=2000,
    flavor="seurat_v3",
    subset=True,
    layer="counts"
)

# The next scvi-tools command allows to determine the main variable to be considered 
# for batch effect correction as well as extra covariates that may add technical 
# noise to downstream analyses. In this case, I consider technology (10X/BD) as 
# as the main batch source, and sample (donor) and library size as a extra covariates.
scvi.data.setup_anndata(
    alldata_sc, batch_key="technology",continuous_covariate_keys=['nFeature_RNA'],
    layer="counts", categorical_covariate_keys=['sample'],
    protein_expression_obsm_key="protein_expression"
)

# Running the totalVI's variational autoencoder (VAE) model
vae_model = scvi.model.TOTALVI(alldata_sc, use_cuda=True, latent_distribution="normal")
vae_model.train()

# The VAE model latent representation has a similar format to a PCA table and can 
# be saved as a matrix that can be used to calculate UMAP in Seurat
# The next file is provided in the data directory
np.savetxt(write_dir+"HealthySamples_BD_all10X_totalVI_latent_representation_byTechBySample.csv", 
vae_model.get_latent_representation(), 
              delimiter=",", fmt='%f')

# Denoised protein and RNA values are calculated while accounting for the 'state' of 
# a cell as represented by the VAE model
# Transformed batch allows to impute protein values as if these were taken from 
# the batch containing observed protein values, in this case, the BD healthy sample
# transformed_batch paramter corresponds to scvi_tools batch metadata (0-indexed)
# as shown in the anndata.obs table 
rna_norm, protein_scaled = vae_model.get_normalized_expression(transform_batch=[1], 
n_samples=25, return_mean=True)

# Denoised feature values are exported for further analysis in Seurat
# The next files are provided in the data directory
protein_scaled.to_csv(write_dir+"HealthySamples_BD_all10X_totalVI_denoised_protein_data_BD_scaled_byTechBySample.tsv", 
sep="\t", float_format="%f")
rna_norm.to_csv(write_dir+"HealthySamples_BD_all10X_totalVI_denoised_rna_data_BD_scaled_byTechBySample.tsv", 
sep="\t", float_format="%f")

```

## Adding totalVI latent space to Seurat

The latent space obtained from totalVI can be used for 
downstream analysis (UMAP/clustering calculation) in Seurat, This is illustrated below.

```{r UMAP/Clustering calculation using totalVI latent space in Seurat, eval=FALSE}

all10X_BD_metadata <- read.table(paste0(data_dir, "all_10xBD_metadata.csv"), 
                                 sep = ",", row.names = 1, header=T)
all10X_BD_metadata <- subset(all10X_BD_metadata, sample %in% 
                                 c("Ctrl_1","CTRL_3", "Healthy"))
allCellNames <- rownames(all10X_BD_metadata)


denoisedProtein <- t(as.matrix(
    read.csv(file = paste0(data_dir, "HealthySamples_BD_all10X_totalVI_denoised_protein_data_BD_scaled_byTechBySample.tsv"),
             sep = "\t", header=T, row.names = 1)))

denoisedRNA <- t(as.matrix(
    read.csv(file= paste0(data_dir, "HealthySamples_BD_all10X_totalVI_denoised_rna_data_BD_scaled_byTechBySample.tsv"),
             sep = "\t", header=T, row.names = 1)))

intData_latent <- as.matrix(
    read.csv(file = paste0(data_dir, "HealthySamples_BD_all10X_totalVI_latent_representation_byTechBySample.csv"), 
             sep = ",", header = F, row.names = allCellNames, 
             colClasses = "numeric"))

colnames(denoisedProtein) <- allCellNames
colnames(denoisedRNA) <- allCellNames

full_denoised_values_matrix <- rbind(denoisedRNA, denoisedProtein)

allHealthies_int_totalVI <- CreateSeuratObject(counts = full_denoised_values_matrix, 
                                               assay = "scvi_denoised")

allHealthies_int_totalVI[['scvi']] <- CreateDimReducObject(embeddings = intData_latent, 
                                                           key = "scvi_", 
                                                           assay = "scvi_denoised")

# UMAP and clustering is performed using the latent space obtained from totalVI
# This reduction replaces PCA, which is the common the standard for this routines.
allHealthies_int_totalVI <- FindNeighbors(allHealthies_int_totalVI,
                                          reduction = 'scvi', dims=1:20)
allHealthies_int_totalVI <- FindClusters(allHealthies_int_totalVI, 
                                         resolution = 0.6, 
                                         graph.name = "scvi_denoised_snn")
allHealthies_int_totalVI <- RunUMAP(allHealthies_int_totalVI, 
                                    reduction = 'scvi', dims=1:20)


tech_metadata_allHealthies <- all10X_BD_metadata$technology
names(tech_metadata_allHealthies) <- rownames(all10X_BD_metadata)
allHealthies_int_totalVI <- AddMetaData(allHealthies_int_totalVI, 
                                               metadata = tech_metadata_allHealthies, 
                                               col.name = "technology")

donor_metadata_allHealthies <- all10X_BD_metadata$sample
names(donor_metadata_allHealthies) <- rownames(all10X_BD_metadata)
allHealthies_int_totalVI <- AddMetaData(allHealthies_int_totalVI, 
                                               metadata = donor_metadata_allHealthies, 
                                               col.name = "donor")

```


## Differential expression analysis

The scvi-tools/totalVI authors do not recommend performing standard differential expression (DE) testing over denoised 
protein/RNA values, as it fails to account for the uncertainity of the denoising correction. In here, I apply DE testing 
over raw feature counts and over the cells where those features where detected (e.g. 10X features over 10X cells) while 
using the unsupervised clustering obtained form the totalVI integrated dataset (BD+10X). Further, I use the Seurat 
'FindConsevedMarkers' function to detect significantly DE genes that are shared between technologies. In the next chunk, 
I apply these operations using a Seurat object containing two assays. The first, "allData", contains the merged raw counts 
matrices for the BD/10X healthy datasets, which are used to calculate DE using the negative binomial test. 
The second, "scvi_denoised", contains the denoised RNA/protein values obtained from the totalVI integration approach. 
Although, as mentioned, these values should not be used in standard DE testing, these can be used 
to display feature expression over UMAP/tSNE projections. Additionally, DE testing with denoised feature counts 
still can render valid results when using the Bayesian pipeline provided by scvi-tools.

Tables of differential expression obtained with the next code are also provided in the data folder. 
To facilitate interpretation of these values, I plot all the per cluster positively DE genes as heatmaps, which are 
available in the plots folder of this repository.


```{r DE testing over BD/10X integrated dataset}

allHealthies_int_totalVI <- readRDS(paste0(data_dir, "allHealthies_int_totalVI_techBatch_pNFeature_SampleCov_final.rds"))

allhealthy_10Xcells <- rownames(subset(allHealthies_int_totalVI@meta.data, technology == "10X"))
allhealthy_BDcells <- rownames(subset(allHealthies_int_totalVI@meta.data, technology == "BD"))

# remove mitochondrial genes and ribosomal subunits 
intFeatures <-  grep(x=rownames(allHealthies_int_totalVI), 
                        pattern = paste(c("^MT-", "^MTRNR", "RPL", "RPS"), 
                                        collapse = "|"), invert = T)

# perform DE testing on raw counts
DefaultAssay(allHealthies_int_totalVI) <- "allData"
# Use clustering obtained with totalVI latent space 
Idents(allHealthies_int_totalVI) <- "scvi_denoised_snn_res.0.6"
levels(allHealthies_int_totalVI) <- c(0:8)

# Markers for 10X cells only
markers_10XOnly <- FindAllMarkers(object = subset(allHealthies_int_totalVI, cells=allhealthy_10Xcells), 
                                  assay = "allData", features = intFeatures, test.use = "negbinom")

# match() - this sets an arrangement for clusters by their putative identity 
# (e.g.: c(0,1,4) == "naive cells"). This arrangement was set 'a posteriori' 
# to DE analysis, but is included in here as it makes expression 
# heatmaps clearer.

top10XMarkers <- markers_10XOnly %>% subset(p_val_adj <= 0.05) %>%
    arrange(match(cluster, c(0,1,4, 6,7, 2,3,5, 8))) %>% 
    group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)

# heatmaps display values in scaled.data slot. For this, raw counts were normalized using 
# standard seurat log normalization. Normalized counts were centred and scaled using 
# Seurat's ScaleData() function while regressing the variables: "nFeature_allData", 
# "percent.mt","S.Score", "G2M.Score", "nCount_allData", and "donor".

top10XFeatures_heatmap <- seurat_expression_heatmap(aSeuratObject = allHealthies_int_totalVI, 
                          specificGenes = unique(subset(top10XMarkers, avg_logFC > 0 )[["gene"]]), 
                          specificIDs = c(0,1,4, 6,7, 2,3,5, 8), 
                          lookAtID = "scvi_denoised_snn_res.0.6",
                          ColPalette = "ggplot",
                          useAssay = "allData", 
                          cells = allhealthy_10Xcells)

top10XFeatures_heatmap <- top10XFeatures_heatmap + 
    labs(title= "Top 10 DE features - unsupervised clustering over totalVI BD/10X integration", 
         subtitle="10X healthy sample cells") + 
  theme(plot.title = element_text(face="bold", size=14), axis.text.y = element_text(size = 12))


# Markers for BD cells only
markers_BDOnly <- FindAllMarkers(object = subset(allHealthies_int_totalVI, cells=allhealthy_BDcells), 
                                  assay = "allData", features = intFeatures, test.use = "negbinom")

topBDMarkers <- markers_BDOnly %>% subset(p_val_adj <= 0.05) %>%
    arrange(match(cluster, c(0,1,4, 6,7, 2,3,5, 8))) %>% 
    group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)


# Next function allows to match colors in the upper heatmap bar with those shown in a 
# UMAP generated using a specific level order. This also subsets the supplied object to 
# idents of interest, in a way that those not requested will not appear empty in 
# the final heatmap, which is the beahvior of the DoHeatmap() original function.
BDtopFeatures_heatmap <- seurat_expression_heatmap(aSeuratObject =  allHealthies_int_totalVI, 
                          specificGenes = unique(subset(topBDMarkers, avg_logFC > 0 )[["gene"]]), 
                          specificIDs = c(0,1,4, 6,7, 2,3,5, 8), 
                          lookAtID = "scvi_denoised_snn_res.0.6",
                          ColPalette = "ggplot",
                          useAssay = "allData", 
                          cells = allhealthy_BDcells)

BDtopFeatures_heatmap <- BDtopFeatures_heatmap + 
    labs(title= "Top 10 DE features - unsupervised clustering over totalVI BD/10X integration", 
         subtitle="BD healthy sample cells") + 
  theme(plot.title = element_text(face="bold", size=14), axis.text.y = element_text(size = 12))

# These parameters are taken by saveThosePlots() to set the final figure size (inches)
global_height <- 9 
global_width <- 13

saveThosePlots(BDtopFeatures_heatmap, 
               aFileName = "BDtopFeatures_heatmap_final.png",
               outDir = write_dir)

saveThosePlots(top10XFeatures_heatmap, 
               aFileName = "top10XFeatures_heatmap_final.png",
               outDir = write_dir)


# Based on protein expression, I merge the naive clusters (0,1,4) in a single cluster
# Remaining clusters are mostly the same, with the exception of EMRA cells, that represent
# a subset from the cluster 5, which cells express CD45RA. Cluster 8 was removed as 
# this cluster is mainly composed of 10X cells only.
# These identities are stored as 'markerCellIDs'.
# I use this clustering to calculate conserved markers.

Idents(allHealthies_int_totalVI) <- "markerCellIDs"

healthies_aggregatedIDs <- c("Naive","Treg naive", "Treg memory", "Effector LGALS1 high", 
                             "Effector KLRB1 high", "Efector cytotoxic", "EMRA")


allHealthies_conserved_markers_2 <- do.call(rbind, lapply(healthies_aggregatedIDs, 
                                                        function(aCellType, 
                                                                 broadIDs=healthies_aggregatedIDs, 
                                                                 aSeuratObject=allHealthies_int_totalVI){
  
  cellTypeBase <- broadIDs[-c(grep(x=broadIDs, pattern = aCellType))]
  
  activated_consMarkers_IDs <- FindConservedMarkers(aSeuratObject, 
                                                    grouping.var = "technology", 
                                                    ident.1 = aCellType, 
                                                    ident.2 = cellTypeBase, 
                                                    slot = "counts", assay="allData", 
                                                    test.use = "negbinom") %>% 
    rownames_to_column("gene") %>% mutate(marker4cellID=aCellType)
  
})
)

# Correlation between avg_expression of BD and 10X cells is similar:
#cor(allHealthies_conserved_markers_2[["BD_avg_logFC"]], allHealthies_conserved_markers_2[["10X_avg_logFC"]])
# [1] 0.9498376
# BD_avg_logFC is thus used to subset to genes with positive expression changes and 
# expression sorting

# Arrange conserved markers based on avg_logFC 

arrangedConservedMarkers <- allHealthies_conserved_markers_2 %>% 
    arrange(match(marker4cellID, healthies_aggregatedIDs)) %>% 
    group_by(marker4cellID) %>% top_n(n = 20, wt = BD_avg_logFC)

ConservedFeatures_heatmap <- seurat_expression_heatmap(aSeuratObject =  allHealthies_int_totalVI, 
                          specificGenes = unique(subset(arrangedConservedMarkers, BD_avg_logFC > 0 )[["gene"]]), 
                          specificIDs = healthies_aggregatedIDs, 
                          lookAtID = "markerCellIDs",
                          ColPalette = "ggsci",
                          useAssay = "allData", label=F)

ConservedFeatures_heatmap <- ConservedFeatures_heatmap + 
    labs(title= "Conserved DE features - marker based cell IDs", 
         subtitle="All cells in the integrated dataset") + 
  theme(plot.title = element_text(face="bold", size=14), axis.text.y = element_text(size = 12))

global_height <- 9 
global_width <- 13

saveThosePlots(ConservedFeatures_heatmap, 
               aFileName = "conservedFeatures_heatmap_final.png",
               outDir = write_dir)

```

# UMAP plots

I the next chunk I provide the code used to generate the remaining figures in the plots folder.

```{r UMAPs for integrated dataset identities}

# remove cells in cluster 8 as these are 10X cells only
cell2Plot <- rownames(subset(allHealthies_int_totalVI@meta.data, !(markerCellIDs == "unknown" )))

Idents(allHealthies_int_totalVI) <- "scvi_denoised_snn_res.0.6"

integration_clustering <- plotA_UMAP(aSeuratObject = allHealthies_int_totalVI, 
           aTitle = "totalVI 10X/BD integration", 
           aSubTitle = "0.6 clustering granularity", pt.size=0.7, label=T, 
           label.box=T, repel=T, cells=cell2Plot) + donwstreamSC_UMAP_theme()

Idents(allHealthies_int_totalVI) <- "technology"

integration_tech <- plotA_UMAP(aSeuratObject = allHealthies_int_totalVI, 
           aTitle = "totalVI 10X/BD integration", 
           aSubTitle = "technology", pt.size=0.7, label=T, 
           label.box=T, repel=T, cells=cell2Plot) + donwstreamSC_UMAP_theme()

Idents(allHealthies_int_totalVI) <- "markerCellIDs"

integration_finalcellIds <- plotA_UMAP(aSeuratObject = allHealthies_int_totalVI, 
           aTitle = "totalVI 10X/BD integration", 
           aSubTitle = "final cell IDs - marker based", pt.size=0.7, label=T, 
           label.box=T, repel=T, cells=cell2Plot, cols=pal_igv()(8)) + donwstreamSC_UMAP_theme()

global_height <- 7
global_width <- 7

saveThosePlots(integration_clustering, 
               aFileName = "integration_clustering_final.png",
               outDir = write_dir)
saveThosePlots(integration_tech, 
               aFileName = "integration_tech_final.png",
               outDir = write_dir)
saveThosePlots(integration_finalcellIds, 
               aFileName = "integration_finalcellIds_final.png",
               outDir = write_dir)

# Plot denoised protein values - e.g CD45RA/R0 
DefaultAssay(allHealthies_int_totalVI) <- "scvi_denoised" # - NOTE!! Feature names in 
# this assay contain points (.) instead of dahses (-)
# Feature values in this slot have not been tranformed

raro_expression_allCells <- annotate_figure(p = ggarrange(
FeaturePlot(allHealthies_int_totalVI, features = c("CD45RA.AHS0029"), 
            pt.size = 0.7, cells = cell2Plot) + featureUMAP_theme(showLegend = "right"),
FeaturePlot(allHealthies_int_totalVI, features = c("CD45RO.AHS0036"), 
            pt.size = 0.7, order=T, cells = cell2Plot) + featureUMAP_theme(showLegend = "right"), 
ncol = 2, nrow=1
), top = text_grob(label = "Imputed protein expression - Integrated 10X/BD datasets", 
                   face="bold", size=13))

global_height <- 4
global_width <- 8
saveThosePlots(raro_expression_allCells, aFileName = "raroExpression_imputed_final.png", outDir = main_dir)


```


